---
title: "Forecasting the 2024 US Presidential Election"
subtitle: "My subtitle if needed"
author: 
  - Tina Kim
  - David Flores
  - Kevin Shao
thanks: "Code and data are available at: https://github.com/DavidFJ207/USPresidentialForecast"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
library(tidyverse)
library(dplyr)
library(knitr)
library(lubridate)
library(ggplot2)
library(ggcorrplot)
library(kableExtra)

```


# Introduction

*Overview paragraph:* The 2024 United States of America Presidential Election is to elect the fourty-seventh president of USA, with the two candidates as Donald Trump, the fourty-fifth president of the United States and representative of the Republican Party, and Kamala Harris, current vice president of the United States and representative of the Democratic Party, after Joe Biden announced that he will not continue participate in this presidential election anymore (@BidenQuits). Democracy can be described as liberal and left-leaning, with the emphasis of economy intervention with implementation of policies such as minimum wage and progressive tax rates. Members of the Republican party tend to follow the free economy, and can be described as right-leaning (@RepDem). The results of the 2024 Presidential Election will determine which political party, Republican or Democratic, will become the leader of the United States for the next four years, largely affecting the policies implemented in the United States during this time period. Due to the strong nation power of the United States, this may even affect the world's economic trends in the four years. Therefore, predicting results of the upcoming 2024 United States Presidential Election is a necessary action.

*Estimand paragraph:* To predict results of the 2024 United States of America Presidential Election, our aim is clearly to conclude which candidate, either Harris or Trump, is more likely to win this election. Our response variable is in turn the probability of one candidate winning the election based on poll data, with our chosen poll as Emerson. When the supporting percentage of one candidate, or the party s/he represents, is only slightly less than 50%, we cannot clearly determine whether s/he would be predicted to have been elected as the fourty-seventh US President or not, since there is a group of people who did not choose to participate in the voting process. Therefore, to increase accuracy of predictions, we will build two models, one model with the estimand as the probability of Harris, or the Democratic party, winning the election, and another with the estimand as the probability of Trump, or the Republican party, winning the election. By comparing the two predictions, one can get a final prediction of who is more likely to be elected as the next President of the United States of America.

*Results paragraph:* Summarize the key findings of the model, highlighting its predictive accuracy and implications.

*Why it matters paragraph:* Explain the broader significance of accurately forecasting election results for politics, society, and policymaking.

*Telegraphing paragraph:* 
The remainder of this paper is structured as follows: Section 2 details the data and measurement process; Section 3 covers model development and results; Section 4 discusses implications and future steps.
The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview
We sourced the "Presidential General Election Polls" dataset from FiveThirtyEight [@FiveThirtyEight2024] and performed an in-depth analysis using the statistical programming language R [@citeR].

Our primary objective was to clean, organize, and analyze U.S. presidential election polling data to provide insights into voter preferences by state. This involved selecting data from reputable pollsters, organizing it by state, and addressing any missing or incomplete entries. Additional datasets were merged where necessary, and after thoroughly addressing missing values, we finalized a polished dataset ready for analysis. Below is a table summarizing the polling data by state, showcasing voter preferences by party and demographic breakdowns. The table highlights key trends and differences across states in terms of party lean and demographic distribution. @fig-predictors-pct-summary provides a snapshot of these findings.

```{r}
#| echo: false
#| warning: false
#| message: false


set.seed(754)
poll_data <- read_csv(here::here(
  "data/02-analysis_data/organized_state_poll_data.csv"))

# Name references
column_refs <- list(
  democrat = "democrat_pct",
  republican = "republican_pct",
  highly_educated = "What is the highest level of education you have attained? College graduate",
  lower_educated = "What is the highest level of education you have attained? High school or less",
  female = "Can you please tell me your gender? Female",
  male = "Can you please tell me your gender? Male",
  non_binary_other = "Can you please tell me your gender? Nonbinary or other",
  caucasian = "For statistical purposes only, can you please tell me your ethnicity? White or Caucasian",
  young_age_18_29 = "What is your age range? 18-29 years",
  young_age_30_39 = "What is your age range? 30-39 years",
  young_age_40_49 = "What is your age range? 40-49 years",
  older_age_50_59 = "What is your age range? 50-59 years",
  older_age_60_69 = "What is your age range? 60-69 years",
  older_age_70_more = "What is your age range? 70 or more years"
)

```

```{r}
#| label: fig-predictors-pct-summary
#| tbl-cap: "Latest Poll Points by Party"
#| echo: false
#| warning: false
#| message: false

poll_summary <- poll_data %>%
  group_by(state) %>%   
  summarise(
    dem_total_pct = sum(democrat_pct, na.rm = TRUE),
    rep_total_pct = sum(republican_pct, na.rm = TRUE),
    highly_educated_pct = mean(`What is the highest level of education you have attained? College graduate`, na.rm = TRUE),
    lower_educated_pct = mean(`What is the highest level of education you have attained? High school or less`, na.rm = TRUE),
    female_pct = mean(`Can you please tell me your gender? Female`, na.rm = TRUE),
    male_pct = mean(`Can you please tell me your gender? Male`, na.rm = TRUE),
    non_binary_other_pct = mean(`Can you please tell me your gender? Nonbinary or other`, na.rm = TRUE),
    caucasian_pct = mean(`For statistical purposes only, can you please tell me your ethnicity? White or Caucasian`, na.rm = TRUE),
    minority_pct = 100 - caucasian_pct,
    age_young_pct = mean(`What is your age range? 18-29 years`, na.rm = TRUE) + 
                    mean(`What is your age range? 30-39 years`, na.rm = TRUE) + 
                    mean(`What is your age range? 40-49 years`, na.rm = TRUE),
    older_pct = mean(`What is your age range? 50-59 years`, na.rm = TRUE) + 
                mean(`What is your age range? 60-69 years`, na.rm = TRUE) + 
                mean(`What is your age range? 70 or more years`, na.rm = TRUE)
  ) %>%
  mutate(across(everything(), ~ replace_na(., 0))) %>%  # Replace NA values with 0
  mutate(across(where(is.numeric), ~ round(., 1)))  # Round to nearest tenth

colnames(poll_summary) <- c("State", "Democrat Pct", "Republican Pct", 
                            "High Educated", "Low Educated", "Female", 
                            "Male", "Nonbinary", "Caucasian", 
                            "Minority Ethnicity", "Young Age", "Older Age")
poll_summary_party <- poll_summary %>%
  select(State, `Democrat Pct`, `Republican Pct`)

poll_summary_demographics <- poll_summary %>%
  select(State, `High Educated`, `Low Educated`, `Female`, 
         `Male`, `Nonbinary`, `Caucasian`, 
         `Minority Ethnicity`, `Young Age`, `Older Age`)

kable(poll_summary_party) %>%
  kable_styling(font_size = 10, position = "center", full_width = FALSE)

# Second table for demographic variables
kable(poll_summary_demographics) %>%
  kable_styling(font_size = 6, position = "center", full_width = FALSE)

```
\newpage  
## Measurement
For this analysis, we selected Emerson as our primary pollster. The rationale behind this choice is explained in detail in @sec-appenA. Emerson’s polls provided essential details, including polling date ranges, party affiliations, sample sizes, and the percentage breakdown of support for different political parties.

We focused on analyzing voter preferences based on party affiliation rather than individual candidate support, as this was more relevant to predicting overall trends. By capturing the broader party dynamics, our model aims to forecast how states are likely to swing in future elections, especially with respect to swing states. This focus will be elaborated in our results section, @sec-results.

Additionally, the Emerson dataset included critical demographic data, such as respondents' political preferences, motivations for voting in the 2024 election, opinions on Donald Trump’s legal battles, and perspectives on major political figures and issues. These details, coupled with information on party affiliation, sources of political news, and other personal attributes, allowed us to paint a richer picture of the voting landscape.

To ensure uniformity across all states, we concentrated on questions that were consistently asked across the entire dataset. This enabled us to perform state-level aggregation and analysis without being limited by state-specific polling questions, ensuring a standardized approach across the board.

## Outcome variables
The initial analysis of the dataset revealed some key trends in party preferences across different states. In @fig-party-results, we illustrate which states exhibit the strongest leanings toward either the Democratic or Republican party. This visualization plays a crucial role in identifying the pivotal swing states that lie close to the center of the graph. These swing states, which show near-equal support for both parties, are of particular interest for future election predictions, as they are likely to determine the overall election outcome.

```{r}
#| label: fig-party-results
#| fig-cap: Example Predictor Variable Visualization
#| echo: false

poll_summary <- poll_data %>%
  group_by(state) %>%   
  summarise(
    dem_total_pct = sum(democrat_pct, na.rm = TRUE),
    rep_total_pct = sum(republican_pct, na.rm = TRUE)
  ) %>%
   mutate(party_diff = dem_total_pct - rep_total_pct)

# Create the vertical bar graph
ggplot(poll_summary, aes(x = party_diff, y = reorder(state, party_diff))) +
  geom_col(aes(fill = party_diff > 0), show.legend = FALSE) +  
  scale_fill_manual(values = c("red", "blue")) +  
  labs(x = "Party Lean (Democrat vs. Republican)", y = "State",
       title = "Party Leaning by State (Democrat vs. Republican)") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 6))  

```

By visualizing the difference between Democratic and Republican preferences across states, we are able to pinpoint which states are firmly partisan and which fall closer to the center, indicating their potential to swing either way in future elections.

## Predictor variables

Next, we take a deeper dive into the various demographic and social factors that influence voter preferences. The predictors we consider include education levels, gender, ethnicity, and age group. These variables are crucial as they shape political preferences and are likely to influence which way a state leans politically.

In @fig-predictors-party-pct, we explore how these factors correlate with party lean across the U.S. states. For instance, higher levels of education may correlate with a greater likelihood of supporting the Democratic party, while other demographic factors such as ethnicity and age may play different roles in shaping political affiliation. This analysis helps us to identify which factors have the greatest influence in each state, offering a clearer understanding of what drives the electorate.

```{r}
#| label: fig-predictors-party-pct
#| fig-cap: Example Predictor Variable Visualization
#| echo: false

# Calculate poll_summary with rounding to the nearest tenth
poll_summary <- poll_data %>%
  group_by(state) %>%
  summarise(
    dem_total_pct = round(sum(.data[[column_refs$democrat]], na.rm = TRUE), 1),
    rep_total_pct = round(sum(.data[[column_refs$republican]], na.rm = TRUE), 1),
    highly_educated_pct = round(mean(.data[[column_refs$highly_educated]], na.rm = TRUE), 1),
    lower_educated_pct = round(mean(.data[[column_refs$lower_educated]], na.rm = TRUE), 1),
    female_pct = round(mean(.data[[column_refs$female]], na.rm = TRUE), 1),
    male_pct = round(mean(.data[[column_refs$male]], na.rm = TRUE), 1),
    non_binary_other_pct = round(mean(.data[[column_refs$non_binary_other]], na.rm = TRUE), 1),
    caucasian_pct = round(mean(.data[[column_refs$caucasian]], na.rm = TRUE), 1),
    minority_pct = round(100 - caucasian_pct, 1),
    age_young_pct = round(mean(.data[[column_refs$young_age_18_29]], na.rm = TRUE) + 
                          mean(.data[[column_refs$young_age_30_39]], na.rm = TRUE) + 
                          mean(.data[[column_refs$young_age_40_49]], na.rm = TRUE), 1),
    older_pct = round(mean(.data[[column_refs$older_age_50_59]], na.rm = TRUE) + 
                      mean(.data[[column_refs$older_age_60_69]], na.rm = TRUE) + 
                      mean(.data[[column_refs$older_age_70_more]], na.rm = TRUE), 1)
  ) %>%
  mutate(party_diff = round(dem_total_pct - rep_total_pct, 1)) %>%
  drop_na()

# Select relevant numeric columns for correlation matrix, excluding PartyDifference
corr_data <- poll_summary %>%
  select(
    Democrat = dem_total_pct,
    Republican = rep_total_pct,
    HighlyEducated = highly_educated_pct,
    LowerEducated = lower_educated_pct,
    Female = female_pct,
    Male = male_pct,
    NonBinary = non_binary_other_pct,
    Caucasian = caucasian_pct,
    Minority = minority_pct,
    YoungAge = age_young_pct,
    OlderAge = older_pct
  )

# Calculate the correlation matrix
cor_matrix <- cor(corr_data, use = "complete.obs")

# Set correlations less than -0.9 to NA
cor_matrix[cor_matrix < -0.9] <- NA

# Visualize the correlation matrix without hierarchical clustering
ggcorrplot(cor_matrix, 
           hc.order = FALSE,  
           type = "lower", 
           lab = TRUE, 
           lab_size = 2,  # Size for correlation labels
           title = "Correlation Matrix of Predictors and Party Lean",
           colors = c("blue", "white", "red"),  
           legend.title = "Correlation") +
  theme(
    plot.title = element_text(size = 12, hjust = 0.5),  # Title size and centered
    axis.text = element_text(size = 1),  # Axis text size
    legend.text = element_text(size = 10),  # Legend text size
    legend.title = element_text(size = 10),  # Legend title size
    axis.title = element_text(size = 1),  # Axis labels size
  )


```
The correlation matrix in @fig-predictors-party-pct provides a detailed look at how various predictors are related to party preferences. From this, we can see which predictors—such as education or gender- are more strongly associated with Democratic or Republican leanings. The blue shading indicates stronger correlations with Democratic support, while red represents stronger correlations with Republican support.

This analysis offers valuable insights into how specific demographic factors shape the political landscape in different states, allowing us to better predict electoral outcomes based on shifting demographic trends.

# Pollster Methodology Overview (Appendix A) {#sec-appenA}  
Append a detailed review of a chosen pollster's methodology, survey techniques, strengths, and weaknesses.

# Model {#sec-model}

## Model Development
The goal of our model is to forecast the popular vote outcome of the 2024 US presidential election, and we will choose the predictor variables based on the the correlation matrix we created previously (@fig-predictors-party-pct). Also, if we choose either the percentage of people who vote for Trump (representing the Republican Party) or Harris (representing the Democratic Party), that number being slightly below 50% does not necessarily means that the corresponding person will not be elected as the next President of the United States, since there is a percentage of people who choose to not vote. Therefore, our final choice is to construct two linear/generalized linear models, one for the percentage that Trump will become the next President and one for Harris respectively. One can then compare the two predicted percentage of people voting for the corresponding President to get the final prediction of who is more likely to be the next President of the United States.

Now we will determine whether to choose generalized linear or linear model for each of the two models by checking the respective normalities of the response variables, which will be the percentage of people supporting the Republican party and the percentage of people supoporting the Democratic party respectively. The following @fig-hist-for-republican is the histogram of the Republican Party Supporting Rate, describing the distribution of the response variable for the first model.

```{r}
#| label: fig-hist-for-republican
#| fig-cap: Distribution of Republican Party Supporting Rate
#| echo: false

# Calculate the variable that represents the Republican Party Supporting Rate
# grouped based on various states
data_rep1 <- poll_data %>%
  group_by(state) %>%
  summarise(
    rep_total_pct = round(sum(.data[[column_refs$republican]], na.rm = TRUE), 1)
  )

# Create a histogram based on targeted data variable
ggplot(data = data_rep1, aes(x=rep_total_pct)) +
  geom_histogram(binwidth = 2) +
  labs(title = "Republic Party Supporting Rate Distirbution", x = "Percentage", y = "Count")

```
And the following @fig-hist-for-democratic is the histogram of the Democratic Party Supporting Rate, describing the distribution of the response variable for the second model.
```{r}
#| label: fig-hist-for-democratic
#| fig-cap: Distribution of Democratic Party Supporting Rate
#| echo: false

# Calculate the variable that represents the Republican Party Supporting Rate
# grouped based on various states
data_rep1 <- poll_data %>%
  group_by(state) %>%
  summarise(
    dem_total_pct = round(sum(.data[[column_refs$democrat]], na.rm = TRUE), 1)
  )

# Create a histogram based on targeted data variable
ggplot(data = data_rep1, aes(x=dem_total_pct)) +
  geom_histogram(binwidth = 2) +
  labs(title = "Democratic Party Supporting Rate Distirbution", x = "Percentage", y = "Count")

```
According to @fig-hist-for-democratic and @fig-hist-for-democratic, both response variables' distributions do not seem to be normal. The distribution of the variable democratic party supporting rate tends to be poisson, since most percentages have a similar frequency, with an outliar at around 67 to 68 percent. The distribution of the variable republic party supporting rate has a maximum frequency at the percentage of around 50 percent, but has similaer frequencies at around 25 to 26 percent and around 48 to 49 percent, which may indicate a violation of the normal distribution property. Therefore, since the response variables of both models are not normally-distributed, a generalized linear model, instead of a linear model, will be used for both models.

*Person B:* Choose a linear or generalized linear model. Justify the selection based on your research goal and data structure.

## Model Set-up
Define the model mathematically and contextually (e.g., linear model predicting vote share).

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i x_i \\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.


*Person B*: Define all model components and assumptions.

## Model Justification
Our results are summarized in @tbl-modelresults. Discuss the implications of the findings and their predictive accuracy.


# Results {#sec-results}

Our results are summarized in @tbl-modelresults.

```{r}
#| label: tbl-modelresults
#| tbl-cap: "Forecasting model results based on poll data"
#| echo: false

# Display results of the model
```
*Person B*: Visualize the model's results and include any performance metrics (e.g., RMSE, test/train split).
```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: ""
#| warning: false

```


# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

Discuss what the model reveals about the election forecast and its potential impact on understanding voting behavior.

*Person C*: Discuss limitations of the model and areas for further improvement.

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

## Appendix A: Pollster Methodology Overview
*Person A*: Provide a detailed methodology review for the chosen pollster, including survey design, population sampling, non-response handling, etc.

## Appendix B: Idealized Survey Design for $100K Budget
*Person C*: Develop and describe an ideal survey design for forecasting the election with a $100K budget, and include survey link.

Sampling approach: The sampling approach we will use is stratified random sampling where we divide the population into strata based on.... (age, gender, education level, etc.)
Recruit Respondents: We will recruit respondents using an online survey on Google Forms so that we can minimize the cost and maximize the range of respondents we can reach. We will spend a portion of our budget (specify here) to advertise these surveys and also send out emails, with an additional monetary incentive (specify here) to encourage more participation. 
Data validation: IP address tracking to prevent duplicate responses.
Poll Aggregation: Incorporate Bayesian Inference.

Survey Link: (Short questionnaire asking for demographic questions to be added here)
Copy of Survey: 

## Additional Data & Model Details
Include any technical details on data cleaning, model diagnostics, and posterior checks.



# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows... 

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

```


\newpage


# References


