---
title: "Forecasting the 2024 US Presidential Election"
subtitle: "My subtitle if needed"
author: 
  - Tina Kim
  - David Flores
  - Kevin Shao
thanks: "Code and data are available at: https://github.com/DavidFJ207/USPresidentialForecast"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```


# Introduction

*Overview paragraph:* Provide a brief overview of the context of the upcoming US presidential election and the importance of forecasting its results.

*Estimand paragraph:* Define the estimand clearly (e.g., predicting the probability of a candidate winning the election based on poll data).

*Results paragraph:* Summarize the key findings of the model, highlighting its predictive accuracy and implications.

*Why it matters paragraph:* Explain the broader significance of accurately forecasting election results for politics, society, and policymaking.

*Telegraphing paragraph:* 
The remainder of this paper is structured as follows: Section 2 details the data and measurement process; Section 3 covers model development and results; Section 4 discusses implications and future steps.
The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview
We use the statistical programming language R [@citeR] and utilize the "Presidential general election polls" dataset [@source]. 

*Person A:* Provide an overview of the dataset's context, sources, and key variables.

## Measurement
Describe how the real-world phenomenon (voter preference) is translated into a dataset entry.

*Person A:* Discuss the survey design, the process of data collection, and potential biases or errors in measurement.

## Outcome variables
Add graphs, tables, and text for each outcome variable. Use subsections as needed.


Some of our data is (@fig-outcome), from @source

```{r}
#| label: fig-outcome
#| fig-cap: Example Outcome Variable Visualization
#| echo: false
# Visualization code

```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it. 

## Predictor variables
Add graphs, tables, and text to describe key predictors.

*Person A:*: Visualize variable distributions (e.g., histograms, boxplots), and justify the inclusion of predictors in the model.

```{r}
#| label: fig-predictors
#| fig-cap: Example Predictor Variable Visualization
#| echo: false
# Visualization code

```

# Pollster Methodology Overview (Appendix A)
Append a detailed review of a chosen pollster's methodology, survey techniques, strengths, and weaknesses.

# Model {#sec-model}

## Model Development
The goal of our model is to forecast the popular vote outcome of the 2024 US presidential election.

*Person B:* Choose a linear or generalized linear model. Justify the selection based on your research goal and data structure.

## Model Set-up
Define the model mathematically and contextually (e.g., linear model predicting vote share).

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i x_i \\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.


*Person B*: Define all model components and assumptions.

## Model Justification
Our results are summarized in @tbl-modelresults. Discuss the implications of the findings and their predictive accuracy.


# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| label: tbl-modelresults
#| tbl-cap: "Forecasting model results based on poll data"
#| echo: false

# Display results of the model
```
*Person B*: Visualize the model's results and include any performance metrics (e.g., RMSE, test/train split).
```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: ""
#| warning: false

```


# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

Discuss what the model reveals about the election forecast and its potential impact on understanding voting behavior.

*Person B*: Discuss limitations of the model and areas for further improvement.

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

## Appendix A: Pollster Methodology Overview
*Person A*: Provide a detailed methodology review for the chosen pollster, including survey design, population sampling, non-response handling, etc.

## Appendix B: Idealized Survey Design for $100K Budget
*Person C*: Develop and describe an ideal survey design for forecasting the election with a $100K budget, and include survey link.

## Additional Data & Model Details
Include any technical details on data cleaning, model diagnostics, and posterior checks.



# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows... 

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```


\newpage


# References


